---
title: "viz_eda_2"
author: "Kaila Boyer"
date: "10/7/2021"
output: github_document
---

```{r include = FALSE}
library(tidyverse)
library(ggridges)
library(viridis) 
library(patchwork)
```

Can include code in the beginning of a document that sets the themes/settings/color for all ggplots 
(look at lecture page on website for code)

Loading in dataset 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

tmax vs. tmin 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs( 
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Date from noaa package with three stations")
```


## Scales 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs( 
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Date from noaa package with three stations") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 C", "0", "15 C")
  ) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
  )
```

## Color Scales 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs( 
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from noaa package with three stations") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 C", "0", "15 C")
  ) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
  ) + 
  scale_color_hue(
    name = "Location", 
    h = c(100,300))

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs( 
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from noaa package with three stations") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 C", "0", "15 C")
  ) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
  ) + 
  scale_color_hue(
    name = "Location") +
  scale_color_viridis_d()
```

## Themes 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs( 
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from noaa package with three stations") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 C", "0", "15 C")
  ) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
  ) + 
  scale_color_hue(
    name = "Location", 
    h = c(100,300)) + 
  scale_color_viridis_d() +
  theme_bw() +
  theme(legend.position = "bottom") 


weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs( 
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from noaa package with three stations") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 C", "0", "15 C")
  ) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
  ) + 
  scale_color_hue(
    name = "Location", 
    h = c(100,300)) + 
  scale_color_viridis_d() +
  theme_classic() +
  theme(legend.position = "bottom") 

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs( 
    title = "Temperature at three stations", 
    x = "Minimum daily temp (C)", 
    y = "Maximum daily temp (C)", 
    caption = "Data from noaa package with three stations" 
    ) + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15 C", "0", "15 C")
  ) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right"
  ) + 
  scale_color_hue(
    name = "Location", 
    h = c(100,300)) + 
  scale_color_viridis_d() +
  theme_minimal() +
  theme(legend.position = "bottom") 
```

## 'data' in geoms 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point()


central_park <- weather_df %>% filter(name == "CentralPark_NY")

waikiki <- weather_df %>%  filter(name == "Waikiki_HA")

waikiki %>% ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point()

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

## 'patchwork' package 

```{r}
ggp_tmax_tmin <- weather_df %>% 
  ggplot(aes(x= tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) +
  theme(legend.position = "none")

ggp_prcp_dens <- weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = 0.3) + 
  theme(legend.position = "none")

ggp_tmax_date <- weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth() + 
  theme(legend.position = "none")

((ggp_tmax_tmin + ggp_prcp_dens) / ggp_tmax_date)
```

## Data Manipulation 

quick example on factors 

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()

weather_df %>% 
  mutate(
    name = fct_reorder(name, tmax)
  ) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

What about tmax and tmin ... 

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin, 
    names_to = "obs", 
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = obs)) + 
  geom_density(alpha = 0.3) + 
  facet_grid(. ~ name)
```
needed to transform the data frame in order to make this kind of plot 


```{r eval = FALSE}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

```







